---
title: "US_census_data"
author: "Emanuel Fitta"
date: "25/5/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

### Importing libraries

```{r}
library(dplyr)
library(readr)
library(tidyr)
```
### Attaching files

```{r}
files <- list.files(pattern = "states_.*csv")
df_list <- lapply(files, read_csv)
us_census <- bind_rows(df_list)
```
### Inspecting 'r us_census' Data Frame

```{r}
head(us_census)
```

```{r}
str(us_census)
```
```{r}
colnames(us_census)
```
### Droping X1 column

```{r}
us_census <- us_census %>%select(-X1)
head(us_census)
```

### Removing $\%$ symbol from each race columns

```{r}
us_census <- us_census %>% 
mutate(Hispanic = gsub('\\%', '', Hispanic), 
White = gsub('\\%','', White),
Black =gsub('\\%','', Black),
Native = gsub('\\%','', Native),
Asian = gsub('\\%', '', Asian),
Pacific = gsub('\\%', '', Pacific), Income = gsub('\\$', '', Income) ) 

head(us_census)
```

### Separating GenderPop column

```{r}
us_census <- us_census %>% 
  separate(GenderPop, into = c("male_pop", "female_pop"), sep = "_")

head(us_census)
```

### Removing M and F from male_pop and female_pop columns

```{r}
us_census <- us_census %>% 
  mutate(male_pop =gsub("\\M", "", male_pop))

us_census <- us_census %>% 
  mutate(female_pop =gsub("\\F", "", female_pop))

head(us_census)
```


### Converting to numeric type

```{r}
us_census <- us_census %>%
  mutate(Hispanic = as.numeric(Hispanic),
         White = as.numeric(White),
         Black = as.numeric(Black),
         Native = as.numeric(Native),
         Asian = as.numeric(Asian),
         Pacific = as.numeric(Pacific),
         Income = as.numeric(Income),
         male_pop = as.numeric(male_pop),
      female_pop = 
        as.numeric(female_pop)
            )

head(us_census)
```

### Converting to decimal the percentages


```{r}
us_census <- us_census %>%
  mutate(Hispanic = Hispanic/100,
         White = White/100,
         Black = Black/100,
         Native = Native/100,
         Asian = Asian/100,
         Pacific = Pacific/100)

us_census
```


### Checking for duplicated rows

```{r}
us_census %>% duplicated()%>% table()

```
### Eliminating duplicated rows

```{r}
us_census <- us_census %>% 
  distinct

us_census %>% duplicated() %>% table()
```
### Gathering 
```{r}
us_census_1 <- us_census %>%
  gather(key = race, value = percentage, "Hispanic":"Pacific")
us_census_1
```
### Hispanic Population

```{r}
hist(us_census$Hispanic, main = "Hispanic Population", xlab = "Hispanic population", col = "green", breaks = 10)
```
```{r}
barplot(height = us_census$Hispanic, names = us_census$State, col = "blue", horiz = F, ylim = c(0,1), main = "Hispanic Population", las = 2, cex.names = 0.5)

```
### White population

```{r}
hist(us_census$White, main = "White Population", xlab = "White population", col = "green", breaks = 10)

```


```{r}
barplot(height = us_census$White, names = us_census$State, col = "blue", horiz = F, ylim = c(0,1), main = "White Population", las = 2, cex.names = 0.5)

```
### Nativepopulation

```{r}
hist(us_census$Native, main = "Native Population", xlab = "Native population", col = "green", breaks = 10, ylim = c(0,50))
```

```{r}
barplot(height = us_census$Native, names = us_census$State, col = "blue", horiz = F, ylim = c(0,0.2), main = "Native Population", las = 2, cex.names = 0.5)

```

### Total population

```{r}
hist(us_census$TotalPop, main = "Total Population", xlab = "Total population")

```

#### Max population

```{r}
max(us_census$TotalPop)
```


```{r}
barplot(height = us_census$TotalPop, names = us_census$State, col = "blue", horiz = F, ylim = c(0,4E07), main = "Total Population", las = 2, cex.names = 0.5)

```

### Male population

```{r}
hist(us_census$male_pop, main = "Male population", xlab = "Male population", ylim = c(0,30))

```

```{r}
barplot(height = us_census$male_pop, names = us_census$State, col = "blue", horiz = F, main = "Male Population", las = 2, cex.names = 0.5, ylim = c(0, 20000000))

```



### Female population

```{r}
hist(us_census$female_pop, main = "Female population", xlab = "Female population", ylim = c(0,30))

```


```{r}
barplot(height = us_census$female_pop, names = us_census$State, col = "blue", horiz = F, main = "Female Population", las = 2, cex.names = 0.5, ylim = c(0, 20000000))

```